Supplementary information

### 2024-04-06 paper/manuscript.Rmd
Update manuscript and supplementary material

# Abstract

Empirical and theoretical evidence suggests that current and projected
environmental changes and biotic homogenisation will dramatically destabilise
ecological communities. However, this evidence is largely based on
oversimplified communities (i.e. food chains, guild assemblages), which lack the
complexity of natural communities.We characterise the dynamics of complex
food-webs under a stochastic bioenergetic food-web model, and investigate how
response diversity, environmental stochasticity, and food-web structure drive
temporal stability. We found that higher average trophic level, omnivory degree
and generalism increased average population stability, but decreased asynchrony,
triggering compensatory dynamics that raised trophic interactions.  Species
richness had little effect on average population stability, but a large one on
synchrony. Further, when response diversity is low, high interaction strength,
average trophic level, and connectance all lead to negative stability-species
richness relationships, but when response diversity is high, we observe positive
stability-species richness relationships regardless of the food-web structure.
Our study reveals that response diversity and food-web structure drive temporal
stability and further improve our understanding of biodiversity-stability
relationships in complex trophic communities.

@@

The ability to do this is supported by a key statistical advance: it is now
possible to partition asynchrony into two component sources: the portfolio
effects (rising from statistical averaging effects, SAE) and the compensatory
dynamics effect (CPE) [@zhao_biodiversity_2022]. Portfolio effects stem from the
statistical averaging of independent population fluctuations [@doak_statistical_1998;
@zhao_biodiversity_2022], thereby increasing asynchrony. Compensatory dynamics
in turn capture population fluctuations that are more synchronous or
asynchronous than expected for independent fluctuations, and can stem from
species interactions and species responses to perturbations [@zhao_biodiversity_2022].
This allows an unprecedented opportunity to obtain mechanistic insight about the
effects of stressors and food-web structure on stability in complex, species
rich communities.

@@

```{r}
met_cap <- "Examples of stochastic food-web simulations. a, Simulations with low
and high response diversity. b, simulations with low and high predator-prey mass
ratio (upper and bottom panels), low and high species richness (left and right
  panels). The food-web are displayed on the side with the node size
being proportional to the average species biomass across the simulation. We
also display the food-web structure and the stability metrics. Across all
simulations, $h = 2$, $\\sigma_e = .3$. Low and high response diversity were
respectively $\\rho = 1$ and $\\rho = 0$. Low and high PPMR were respectively 1 and 100." %>%
  str_replace_all("\\s+", " ")
```

@@

(ref:eq-rho) (eq. \@ref(eq:rho))

```{r}
tar_read(tab_sim_param) %>%
  kable(
    booktabs = TRUE,
    caption = "Parameter values used in the simulation experiment. Numbers
    between squared brackets display all the values. Two numbers between
    brackets are the minimum and maximal values, N being the number of unique
    values. PPMR: Predator Prey Mass Ratio. We defined response diversity as \\protect$1-\\rho$
    (ref:eq-rho).",
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

We generated food-web structure using the niche model [@williams_simple_2000]
with an initial species richness from $`r pS["min"]`$ to $`r pS["max"]`$, and
connectance from $`r pct["min"]`$ to $`r pct["max"]`$ (Table
\@ref(tab:param)). We generated food-webs without cannibalistic links and
discarded food-webs that contained disconnected species.

@@

We numerically solved the stochastic differential equation systems during 2000
timesteps and collected the last 500 timesteps, similarly to previous studies
[@brose_allometric_2006]. We set $d_t = .1$, and $d_t = .05$ when instability
was detected during the simulations due to stiff changes in biomass. We set
extinction threshold to $10^{-6}$ and if an extinction happened in the last 500
timesteps, we continued to run the simulation for another 1000 timesteps until
there was no extinction events during the last 500 timesteps. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we set their biomass to 0 and
re-run the simulation for another 1000 timesteps. Except in the case of
disconnected species, we re-ran simulations with species starting biomass equals
to their biomass at the last timestep of the previous simulation. Supplementary
analysis showed that alternative choices of simulation processes, such as longer
simulations, rebuilding food-webs (i.e. resetting consumer preferences,
recomputing species bodymass according to the new food-web) after removing
disconnected species did not affect our results (Fig. S1). The food-web
generation and simulations were done using the `Julia` package
`EcologicalNetworkDynamics.jl` [@lajaaiti_ecologicalnetworksdynamicsjl_2024].

@@

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
portfolio effects and compensatory dynamic effects). We tested the
effects of food-web structure (average interaction strength, average trophic
level, connectance) on population stability and asynchrony partitions (portfolio and
compensatory effects). Grounded in theoretical ecology, we expected that higher
average trophic level lower interaction strength, lower connectance[@mccann_weak_1998;
@mccann_reevaluating_1997; @shanafelt_stability_2018]. We further added
simulation parameters as control variables because higher Predator-Prey Mass
Ratio and predator interference were also expected to increase population
stability [@brose_allometric_2006]. Finally, we tested the effects of environmental
stochasticity and response diversity on population stability and asynchrony
partitions, although there were predicted to mainly affect to population stability
(i.e. by increasing amplitude of population fluctuations) and asynchrony (i.e.
by generating independent population fluctuations) respectively. We
completed the structural equation model with the mathematical relationship
linking compensatory dynamics and statistical averaging effects to asynchrony,
and from asynchrony and population stability to community stability. We further
developed a more detailled SEM including the effects omnivory and partition of
compensatory dynamics due to species interactions ($CPE_{int}$) and due to
environmental fluctuations ($CPE_{env}$).

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)). We ensured that all the linear models composing the structural
equation model presented low multicollinearity (Variance Inflation
Factor < 3, Table S1), altough interaction strength and species richness were
highly correlated (Fig. S2). The structural equation model using the R package
`PiecewiseSEM` [@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the
SEM to obtain confidence interval on the estimated effects and computed the sum
of the direct and indirect effects using `semEff` R package
[@murphy_semeff_2021]. In the main text, we reported standardised coefficients
which were obtained by scaling the coefficients by the standard deviation of the
response and predictor variable. Finally, we reported only the direct
standardised coefficients that had an absolute value above $0.05$, because
almost all linked were statistically significant given the high numbers of
simulations.

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We then modelled temporal
stability of community biomass according to food-web structure: species
richness, weighted average trophic level, connectance and average interaction
strength, as well as environmental stochasticity and species response diversity.
We further included the two-way interactions between species richness and
food-web structure, plus between species richness and stressors. Finally, we
added the three-way interaction between species richness, food-web structure and
stressors. We added predator interference, predator-prey mass ratio as control
variables. We further added the id of the food-web as a random intercept.  We
then used the linear equation of the model and associated slope coefficients to
predict the shape of the relationship between community stability and species
richness according to a gradient of response diversity and of food-web
structure. To do so, we summed all the coefficients involving species richness
in the model, i.e. the one, two, and three-way terms. The values of the other
variables were set to the values indicated in the Fig. \@ref(fig:pred). The VIF
were inferior to 3, indicating low multicollinearity (Table S2), we checked the
distribution of the residuals (Fig. S3) using `DHARMa` R package. The model was
implemented using the R package `glmmTMB` [@brooks_glmmtmb_2017].

@@

```{r}
sem_cap <- paste0(
  "Structural Equation Model food-web struture and stressors to temporal
  stability of community biomass. (a) Continuous red arrows display positive and
  dashed blue arrows display negative standardised effects, width of the arrows
  being proportional to the absolute values of the standardised effects. Only
  the standardised coefficients whose absolute values are superior to 0.05 are
  displayed. The full table of coefficients including control variables is
  displayed in Table S3. A more complete SEM containing further partition of
  compensatory effects and control variables is displayed in Table S4. (b) Total
  effect of food-web structure on community stability, asynchrony, and
  population stability. N = ", nrow(tar_read(sim_fw)), " simulations.") %>%
  str_replace_all("\\s+", " ")
```

@@

We found that food-web structure had significant effect on temporal population
stability such as average trophic level increase population stability (resp.
$r_\delta$= `r rpt`, $r_\delta$ being the standardised coefficient,
Fig. \@ref(fig:sem)a). In turn, the variance of environmental stochasticity had
a strong negative effect on population stability ($r_\delta$= `r rpe`). We
further found that response diversity had a small negative effect on population
stability ($r_\delta$= `r rpd`).

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and species response
diversity to environmental fluctuations had by far the most important effects on
community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`, Fig.
\@ref(fig:sem)b). Species richness had a overall positive total effect on
community stability ($r_\delta$=`r rtsr`), steming from its strong total positive effect on
asynchrony ($r_\delta$=`r rtar`) and downplayed by its small total negative effect on population
stability ($r_\delta$=`r rtpr`), but the fact that population stability effect on community
stability was twice higher than the one of asynchrony. Average trophic level and
average interaction strength had small positive effects on community stability in
total (resp. $r_\delta$=`r rtst`, `r rtsi`, Fig. \@ref(fig:sem)b), driven by their positive
effects on population stability (resp. $r_\delta$=`r rtpt`, `r rtsi`). However,
we found that average trophic level had a negative total effect on asynchrony
($r_\delta$=`r rtat`), average interaction strength had a positive total effect
($r_\delta$=`r rtai`). In turn, connectance had a small negative total effect on
community stability ($r_\delta$=`r rtsct`), raising from a negative total effect
on population stability ($r_\delta$=`r rtpct`) and a positive one on asynchrony
($r_\delta$=`r rtact`, Fig. \@ref(fig:sem)b).

@@

One of the key question in Ecology is what drive temporal stability in complex
communities?, especially in food-webs. We provide an answer to that question by
investigating how the structure of complex food-webs affects the temporal
stability of ecological communities using a bioenergetic model. We added
environmental stochasticity to this model and modulated the species response
diversity, i.e. the capacity of species to answer differently to environmental
fluctuations. We found that that food-web structure, apart from species richness,
had little effects on overall stability. However we found that the interplay
between food-web structure and response diversity was critical to determine
positive or negative relationships between community stability and species
richness.

@@

We further found that average trophic level increased average population
stability, confirming previous theoretical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due to
lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018]. Additional simulations with equal mortality
rates for all species suggest a prevalence of the former explanation (Fig. S1).
In turn, we found that average trophic level decreases strongly asynchrony
through a negative effect on portfolio effects, by decreasing the evenness of
biomass variability. It is linked to the model assumption that species of higher
trophic levels have lower overall mortality rates, similarly to previous
studies [@vasseur_environmental_2007], thereby increasing average population
stability and simultaneously synchronising populations. Furthermore, average
omnivory was strongly correlated with average trophic level (Fig. S2),
suggesting that average omnivory also stabilise population stability, in line
with previous theoretical results in small modules [@mccann_reevaluating_1997].
Our results further suggest that higher average trophic level results in lower
compensatory dynamics raising from species interactions, this results could be
related by the fact that average trophic level is strongly associated with
average omnivory, which might tamper synchronized predator-prey oscillations
that can raise compensatory dynamics [@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. However, species richness
had a strong positive on asynchrony, by raising portfolio effects.
This result is common in empirical systems [@tilman_biodiversity_2006;
@olivier_urbanization_2020; @zhao_biodiversity_2022], but strongly depends on
the presence of response diversity [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Connectance, quantifying
the generalism in food-webs, had small negative effects population
stability but small positive on asynchrony. We further found that higher connectance
leaded to less even biomass variability, thereby limiting the raise of the
portfolio effects, and leaded to lower compensatory dynamics due to species
interactions (Table. S4). This result is in accordance with empirical findings
that a generalist predator can indeed synchronises the dynamics of a large
number of preys [@raimondo_interspecific_2004], thereby potentially limiting compensatory
dynamics raising from species interactions. However, the overall effect of
connectance on overall stability was quite low, similarly to previous empirical
findings [@danet_species_2021], but in apparent disagreement with the
predominant role of connectance on stability in theoretical studies
[complexity-stability paradox, @may_will_1972].

@@

Predator interference, taken in this study as a control variable, had also a
strong effects on community stability. Our results confirm the stabilising
effect of self-limiting processes on populations [@uszko_effects_2017;
@brose_allometric_2006]. Furthermore, we found that predator interference
decreased asynchrony by strongly decreasing compensatory effects due to species
interactions. Predator interference models a decrease of the feeding rates of
consumers when predator biomass increases which might lead to a less dramatic
fluctuation of prey biomass. All toghether, this result suggest that
self-limitation mechanisms that increase population stability might in turn have
opposite effect on asynchrony, and that it is the balance between the two
effects weighted by the relative importance of population stability and
asynchrony that will determine the overall effect on community stability.

@@

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather control the overall community
stability (i.e. the intercept of the relation). At the opposite, variation in
response diversity had a strong impact on the relationship between stability and
richness. The reason is that the strength of environmental stochasticity mainly
affect population stability and less asynchrony, asynchrony being the driver of
species-richness relationships [@ives_stability_2000; @tilman_biodiversity_2006; @loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Average interaction
strength and connectance that were found to have little effects on overall
stability, were found to have a strong effect on stability-richness
relationships. This result is in accordance with the complexity-stability theory
that predicts that community stability (i.e. measured as the dominant eigen
value of the jacobian matrix) should decrease as the community is species
diverse, has higher connectance, and higher interaction strength
[@may_will_1972]. However, we found that the negative effects of complexity on
stability-richness relationship are cancelled when response diversity is high.
An possible explanation is the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationships. Response diversity might then be
one mechanism that explains why positive relationship between stability
and species richness are so prevalent in empirical settings while food-web
theoretical model that do not often include environmental stochasticity and
response diversity finds negative complexity stability relationships
[@ives_stability_2007].

@@

Our study presents the effects of food-web structure on temporal community
stability and its partition into population stability and asynchrony in complex
communities, added to two pressing stressors in the context of global change,
the strength of environmental variability and the level of response diversity of
species facing this environmental variability. Our results confirm the results
of previous theoretical models modelling small food-web modules or food-chains,
as we report that average trophic level, omnivory, predator-prey size ratio, and
predator interference increase temporal community stability by stabilising
population fluctuations. In turn, our study confirm that species richness
increase temporal community stability by increasing asynchrony due to portfolio
effects and thanks to response diversity. Our study reports that connectance and
average interaction strength have little effects on overall community stability
while they have a strong effect on the sign of the community stability-richness
relationship, in accordance with the complexity-stability paradox, modulated by
response diversity. Overall, it suggests that high response diversity of species
facing environmental stochasticity, on top of being the essential ingredient
driving positive stability-richness relationship, also exceeds the negative
effects of food-web complexity on stability-richness relationship.

Our results highlight that the strength of environmental fluctuations is the
primary driver of community stability. In the context of global changes in which
environmental variability is expected to increase tremendously, our study
suggests a strong decrease in community stability, despite high response
diversity and response diversity. But ongoing biotic homogenization
[@olden_homogocene_2018; @muthukrishnan_invasive_2020; @wang_biotic_2021] is
also suggested to decrease response diversity [@wang_biotic_2021;
@su_human_2021], which in turn is expected to further decrease stability despite
high species richness, even more in complex food-webs having high interaction
strength, high connectance and high average trophic level. Then, the results of
our study reinforce that the idea that restoration of ecosystems along with a
restoration of landscape complexity that provides refuges from extreme
environmental fluctuations will be key to maintain temporal stability of
ecological communities in the coming decades [@ryser_landscape_2021].

>>>>

# Abstract (150/150 words)

Empirical and theoretical evidence suggests environmental changes will
dramatically destabilise ecological communities. However, this evidence is
largely based on oversimplified communities, which lack the complexity of
natural communities. We characterise the dynamics of complex food-webs under a
stochastic food-web model, and investigate how response diversity,
stochasticity, and food-web structure drive temporal stability. We found that
higher average trophic level increased community stability by increasing
population stability, at the expense of asynchrony. Species richness had little
effect on population stability, but a large one on asynchrony. Interaction
strength and connectance had little effects on temporal stability overall, but
interacted strongly with response diversity to determine stability-richness
relationships. Absence of response diversity all leaded to negative
stability-richness relationships, but high response diversity resulted in high
prevalence of positive stability-richness relationships. Our study reveals that
response diversity and food-web structure drive temporal stability overall, and
together determine the sign of the stability-richness relationship.

@@

The ability to do this is supported by a key statistical advance: it is now
possible to partition asynchrony into two component sources: the portfolio
and the compensatory dynamics effects [@zhao_biodiversity_2022]. Portfolio
effects stem from the statistical averaging of independent population
fluctuations [@doak_statistical_1998; @zhao_biodiversity_2022], thereby
increasing asynchrony. Compensatory dynamics in turn capture population
fluctuations that are more synchronous or asynchronous than expected for
independent fluctuations, and can stem from species interactions and species
responses to perturbations [@zhao_biodiversity_2022]. This allows an
unprecedented opportunity to obtain mechanistic insight about the effects of
stressors and food-web structure on stability in complex, species rich
communities.

@@

```{r}
met_cap <- "Examples of stochastic food-web simulations. a, Simulations with low
and high response diversity. b, simulations with low and high predator-prey mass
ratio (upper and bottom panels), low and high species richness (left and right
  panels). The food-webs are displayed on the side with the node size
being proportional to the average species biomass across the simulation. We
also display the food-web structure and the stability metrics. Across all
simulations, $h = 2$, $\\sigma_e = .3$. Low and high response diversity were
respectively $\\rho = 1$ and $\\rho = 0$. Low and high PPMR were respectively 1
and 100." %>%
str_replace_all("\\s+", " ")
```

@@

We generated food-web structure using the niche model [@williams_simple_2000]
with an initial species richness from $`r pS["min"]`$ to $`r pS["max"]`$, and
connectance from $`r pct["min"]`$ to $`r pct["max"]`$ (Table S1). We generated
food-webs without cannibalistic links and discarded food-webs that contained
disconnected species.

@@

We numerically solved the stochastic differential equation systems during 2000
timesteps and collected the last 500 timesteps, similarly to previous studies
[@brose_allometric_2006]. We set $d_t = .1$, and $d_t = .05$ when instability
was detected during the simulations due to stiff changes in biomass. We set
extinction threshold to $10^{-6}$ and if an extinction happened in the last 500
timesteps, we continued to run the simulation for another 1000 timesteps until
there was no extinction events during the last 500 timesteps. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we set their biomass to 0 and
re-run the simulation for another 1000 timesteps. Except in the case of
disconnected species, we re-ran simulations with species starting biomass equal
to their biomass at the last timestep of the previous simulation. Supplementary
analysis showed that alternative choices of simulation processes, such as longer
simulations, rebuilding food-webs (i.e. resetting consumer preferences,
recomputing species bodymass according to the new food-web) after removing
disconnected species did not affect our results (Fig. S1). The food-web
generation and simulations were done using the `Julia` package
`EcologicalNetworkDynamics.jl` [@lajaaiti_ecologicalnetworksdynamicsjl_2024].

@@

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
portfolio effects and compensatory dynamic effects). We tested the
effects of food-web structure (average interaction strength, average trophic
level, connectance) on population stability and asynchrony partitions (portfolio and
compensatory effects). Grounded in theoretical ecology, we expected that higher
average trophic level, lower interaction strength, lower connectance[@mccann_weak_1998;
@mccann_reevaluating_1997; @shanafelt_stability_2018]. We further added
simulation parameters as control variables because higher Predator-Prey Mass
Ratio and predator interference were also expected to increase population
stability [@brose_allometric_2006]. Finally, we tested the effects of environmental
stochasticity and response diversity on population stability and asynchrony
partitions, although there were predicted to mainly affect to population stability
(i.e. by increasing amplitude of population fluctuations) and asynchrony (i.e.
by generating independent population fluctuations) respectively. We
completed the structural equation model with the mathematical relationship
linking compensatory dynamics and statistical averaging effects to asynchrony,
and from asynchrony and population stability to community stability. We further
developed a more detailled SEM including the effects omnivory and partition of
compensatory dynamics due to species interactions ($CPE_{int}$) and due to
environmental fluctuations ($CPE_{env}$).

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)). We ensured that all the linear models composing the structural
equation model presented low multicollinearity (Variance Inflation
Factor < 3, Table S2), although interaction strength and species richness were
highly correlated (Fig. S2). The structural equation model using the R package
`PiecewiseSEM` [@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the
SEM to obtain confidence intervals on the estimated effects and computed the sum
of the direct and indirect effects using `semEff` R package
[@murphy_semeff_2021]. In the main text, we reported standardised coefficients
which were obtained by scaling the coefficients by the standard deviation of the
response and predictor variable. Finally, we reported only the direct
standardised coefficients that had an absolute value above $0.05$, because
almost all linked were statistically significant given the high numbers of
simulations.

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We then modelled temporal
stability of community biomass according to food-web structure: species
richness, weighted average trophic level, connectance and average interaction
strength, as well as environmental stochasticity and species response diversity.
We further included the two-way interactions between species richness and
food-web structure, plus between species richness and stressors. Finally, we
added the three-way interaction between species richness, food-web structure and
stressors. We added predator interference, predator-prey mass ratio as control
variables. We further added the id of the food-web as a random intercept.  We
then used the linear equation of the model and associated slope coefficients to
predict the shape of the relationship between community stability and species
richness according to a gradient of response diversity and of food-web
structure. To do so, we summed all the coefficients involving species richness
in the model, i.e. the one, two, and three-way terms. The values of the other
variables were set to the values indicated in the Fig. \@ref(fig:pred). The VIF
were inferior to 3, indicating low multicollinearity (Table S3), we checked the
distribution of the residuals (Fig. S3) using `DHARMa` R package. The model was
implemented using the R package `glmmTMB` [@brooks_glmmtmb_2017].

@@

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking food-web structure and stressors to temporal
  stability of community biomass. a, continuous red arrows display positive and
  dashed blue arrows display negative standardised effects, width of the arrows
  being proportional to the absolute values of the standardised effects. Only
  the standardised coefficients whose absolute values are superior to 0.05 are
  displayed. The full table of coefficients including control variables is
  displayed in Table S4. A more complete SEM containing further partition of
  compensatory effects and control variables is displayed in Table S5. b, total
  effects of food-web structure on community stability, asynchrony, and
  population stability. N = ", nrow(tar_read(sim_fw)), " simulations.") %>%
  str_replace_all("\\s+", " ")
```

@@

We found that food-web structure had significant effects on temporal population
stability such as average trophic level increase population stability (resp.
$r_\delta$= `r rpt`, $r_\delta$ being the standardised coefficient,
Fig. \@ref(fig:sem)a). In turn, the variance of environmental stochasticity had
a strong negative effect on population stability ($r_\delta$= `r rpe`). We
further found that response diversity had a small negative effect on population
stability ($r_\delta$= `r rpd`).

@@

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and species response
diversity to environmental fluctuations had by far the most important effects on
community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`, Fig.
\@ref(fig:sem)b). Species richness had a overall positive total effect on
community stability ($r_\delta$=`r rtsr`), stemming from its strong total positive effect on
asynchrony ($r_\delta$=`r rtar`) and downplayed by its small total negative effect on population
stability ($r_\delta$=`r rtpr`), but population stability effect on community
stability was twice higher than the one of asynchrony. Average trophic level and
average interaction strength had small positive effects on community stability in
total (resp. $r_\delta$=`r rtst`, `r rtsi`, Fig. \@ref(fig:sem)b), driven by their positive
effects on population stability (resp. $r_\delta$=`r rtpt`, `r rtsi`). However,
we found that average trophic level had a negative total effect on asynchrony
($r_\delta$=`r rtat`), average interaction strength had a positive total effect
($r_\delta$=`r rtai`). In turn, connectance had a small negative total effect on
community stability ($r_\delta$=`r rtsct`), raising from a negative total effect
on population stability ($r_\delta$=`r rtpct`) and a positive one on asynchrony
($r_\delta$=`r rtact`, Fig. \@ref(fig:sem)b).

@@

One of the key questions in Ecology is what drives temporal stability in complex
communities?, especially in food-webs. We provide an answer to that question by
investigating how the structure of complex food-webs affects the temporal
stability of ecological communities using a bioenergetic model. We added
environmental stochasticity to this model and modulated the species response
diversity, i.e. the capacity of species to answer differently to environmental
fluctuations. We found that food-web structure, apart from species richness,
had little effects on overall stability. However we found that the interplay
between food-web structure and response diversity was critical to determine
positive or negative relationships between community stability and species
richness.

@@

We further found that average trophic level increased average population
stability, confirming previous theoretical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due to
lower mortality rates which in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018]. Additional simulations with equal mortality
rates for all species suggest a prevalence of the former explanation (Fig. S1).
In turn, we found that average trophic level decreases strongly asynchrony
through a negative effect on portfolio effects, by decreasing the evenness of
biomass variability. It is linked to the model assumption that species of higher
trophic levels have lower overall mortality rates, similarly to previous
studies [@vasseur_environmental_2007], thereby increasing average population
stability and simultaneously synchronising populations. Furthermore, average
omnivory was strongly correlated with average trophic level (Fig. S2),
suggesting that average omnivory also stabilises population stability, in line
with previous theoretical results in small modules [@mccann_reevaluating_1997].
Our results further suggest that higher average trophic level results in lower
compensatory dynamics raising from species interactions, this results could be
related by the fact that average trophic level is strongly associated with
average omnivory, which might tamper synchronised predator-prey oscillations
that can raise compensatory dynamics [@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. However, species richness
had a strong positive effect on asynchrony, by raising portfolio effects.
This result is common in empirical systems [@tilman_biodiversity_2006;
@olivier_urbanization_2020; @zhao_biodiversity_2022], but strongly depends on
the presence of response diversity [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Connectance, quantifying
the generalism in food-webs, had small negative effects on population
stability but a small positive on on asynchrony. We further found that higher connectance
led to less even biomass variability, thereby limiting the rise of the
portfolio effects, and led to lower compensatory dynamics due to species
interactions (Table. S4). This result is in accordance with empirical findings
that a generalist predator can indeed synchronise the dynamics of a large
number of prey [@raimondo_interspecific_2004], thereby potentially limiting compensatory
dynamics arising from species interactions. However, the overall effect of
connectance on overall stability was quite low, similarly to previous empirical
findings [@danet_species_2021], but in apparent disagreement with the
predominant role of connectance on stability in theoretical studies
[complexity-stability paradox, @may_will_1972].

@@

Predator interference, taken in this study as a control variable, had also a
strong effect on community stability. Our results confirm the stabilising
effect of self-limiting processes on populations [@uszko_effects_2017;
@brose_allometric_2006]. Furthermore, we found that predator interference
decreased asynchrony by strongly decreasing compensatory effects due to species
interactions. Predator interference models a decrease of the feeding rates of
consumers when predator biomass increases which might lead to a less dramatic
fluctuation of prey biomass. All together, this result suggests that
self-limitation mechanisms that increase population stability might in turn have
opposite effects on asynchrony, and that it is the balance between the two
effects weighted by the relative importance of population stability and
asynchrony that will determine the overall effect on community stability.

@@

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather controls the overall community
stability (i.e. the intercept of the relation). At the opposite, variation in
response diversity had a strong impact on the relationship between stability and
richness. The reason is that the strength of environmental stochasticity mainly
affect population stability and less asynchrony, asynchrony being the driver of
species-richness relationships [@ives_stability_2000; @tilman_biodiversity_2006; @loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Average interaction
strength and connectance that were found to have little effects on overall
stability, were found to have a strong effect on stability-richness
relationships. This result is in accordance with the complexity-stability theory
that predicts that community stability (i.e. measured as the dominant eigen
value of the jacobian matrix) should decrease as the community is species
diverse, has higher connectance, and higher interaction strength
[@may_will_1972]. However, we found that the negative effects of complexity on
stability-richness relationships are cancelled when response diversity is high.
A possible explanation is the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationships. Response diversity might then be
one mechanism that explains why positive relationships between stability
and species richness are so prevalent in empirical settings while food-web
theoretical models that do not often include environmental stochasticity and
response diversity finds negative complexity stability relationships
[@ives_stability_2007].

@@

Our study presents the effects of food-web structure on temporal community
stability and its partition into population stability and asynchrony in complex
communities, added to two pressing stressors in the context of global change,
the strength of environmental variability and the level of response diversity of
species facing this environmental variability. Our results confirm the results
of previous theoretical models modelling small food-web modules or food-chains,
as we report that average trophic level, omnivory, predator-prey size ratio, and
predator interference increase temporal community stability by stabilising
population fluctuations. In turn, our study confirms that species richness
increases temporal community stability by increasing asynchrony due to portfolio
effects and thanks to response diversity. Our study reports that connectance and
average interaction strength have little effects on overall community stability
while they have a strong effect on the sign of the community stability-richness
relationship, in accordance with the complexity-stability paradox, modulated by
response diversity. Overall, it suggests that high response diversity of species
facing environmental stochasticity, on top of being the essential ingredient
driving positive stability-richness relationship, also exceeds the negative
effects of food-web complexity on stability-richness relationship.

Our results highlight that the strength of environmental fluctuations is the
primary driver of community stability. In the context of global changes in which
environmental variability is expected to increase tremendously, our study
suggests a strong decrease in community stability, despite high response
diversity and response diversity. But ongoing biotic homogenization
[@olden_homogocene_2018; @muthukrishnan_invasive_2020; @wang_biotic_2021] is
also suggested to decrease response diversity [@wang_biotic_2021;
@su_human_2021], which in turn is expected to further decrease stability despite
high species richness, even more in complex food-webs having high interaction
strength, high connectance and high average trophic level. Then, the results of
our study reinforce the idea that restoration of ecosystems along with a
restoration of landscape complexity that provides refuges from extreme
environmental fluctuations will be key to maintain temporal stability of
ecological communities in the coming decades [@ryser_landscape_2021].

### 2024-04-06 paper/supplementary_figures.Rmd
Update manuscript and supplementary material

# Table

```{r}
tar_read(tab_sem_vif) %>%
  kable(booktabs = TRUE,
    format = "latex",
    caption = "Variance Inflation Factors for the Structural Equation Model
    displayed in main text (Fig. 2). All VIF were inferior to 3, indicating low
    multicollinearity.",
    label = "sem-vif"
    ) %>%
  column_spec(1, bold = TRUE) %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

>>>>

# Tables

```{r}
tar_load(param_comb_nb)
param_comb_chr <- paste0("$",
  paste0(latex_symbol_replacement()[colnames(param_comb_nb)[colnames(param_comb_nb) != "n"]], collapse = " \\times "),
  " =$ ", nrow(param_comb_nb), " parameter combinations."
)
nb_fw_per_comb <- unique(param_comb_nb$n)

param_table_cap <- paste0(
  "Parameter values used in the simulation experiment. N = ",
  nrow(tar_read(sim_fw)), " simulations. ", param_comb_chr,
  " We generated ", nb_fw_per_comb, " food-webs of varying connectance for each
  parameter combination.") %>%
  str_replace_all("\\s+", " ")

tar_read(param_table) %>%
  kable(
    booktabs = TRUE,
    format = "latex",
    escape = FALSE,
    caption = param_table_cap,
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
tar_read(tab_sem_vif) %>%
  kable(
    booktabs = TRUE,
    format = "latex",
    caption = "Variance Inflation Factors for the Structural Equation Model
    displayed in main text (Fig. 2). All VIF were inferior to 3, indicating low
    multicollinearity.",
    label = "sem-vif"
    ) %>%
  column_spec(1, bold = TRUE) %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```
